---
title       : Fighting Information Overload 
subtitle    : with Word Clouds
author      : Knut Behrends, Sept. 2014
logo        : Screenshot-shinyapp-wordclouds.png
job         : Reproducible Pitch Presentation for Online Course 'Developing Data Products'
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---



## Slide 2
### Information Overload in the STEM Sciences

- There is more scientific literature produced than can be possibly read
- A staggering 6,000 papers are published every day 
- Source: [Nature News & Comment, Sept. 3, 2014](http://www.nature.com/news/how-to-tame-the-flood-of-literature-1.15806)

###  Information Overload in small subfields of sciences

- Even a single big conference releases an overwhelming amount of information
- Let's develop a new tool to tame this
- Stategy: Interactively summarize the summaries

--- 

## Slide 3
### My Shiny App: Word Clouds

- it visualizes the most common phrases used in conference abstracts, AGU Fall Meeting 2012
- Each word-cloud picture "summarizes the summaries" of hundreds of science talks and posters.

- Example: 



--- 

## Slide 4

### The American Geophysical Union (AGU) 

- is a professional society for Earth Scientists. 
-  Once a year, there is a big conference with more than 20000 participants from around the world.

- Over the years, number of participants has increased considerably.
```{r echo=F, return="hide"}
library(ggplot2)
library(rjson)
options(warn=-1)
df <- rjson::fromJSON("{\"year\":[\"2005\",\"2006\",\"2007\",\"2008\",\"2009\",\"2010\",\"2011\",\"2012\",\"2004\",\"2005\",\"2006\",\"2007\",\"2008\",\"2009\",\"2010\",\"2011\",\"2012\",\"2013\"],\"participants\":[\"8067\",\"7732\",\"8037\",\"8685\",\"9088\",\"10463\",\"10725\",\"11275\",\"11422\",\"12008\",\"13972\",\"14649\",\"16242\",\"16642\",\"19175\",\"22167\",\"24000\",\"22500\"],\"orals\":[\"NA\",\"NA\",\"NA\",\"NA\",\"NA\",\"4431\",\"4333\",\"4436\",\"NA\",\"NA\",\"NA\",\"NA\",\"NA\",\"NA\",\"5852\",\"6000\",\"6800\",\"7500\"],\"posters\":[\"NA\",\"NA\",\"NA\",\"NA\",\"NA\",\"9370\",\"8439\",\"9092\",\"NA\",\"15788\",\"NA\",\"NA\",\"15730\",\"NA\",\"11517\",\"12000\",\"14000\",\"15000\"],\"conference\":[\"EGU\",\"EGU\",\"EGU\",\"EGU\",\"EGU\",\"EGU\",\"EGU\",\"EGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\",\"AGU\"]}")
df <- as.data.frame(df)
df <- df[df$conference=="AGU", ]
df$participants <- as.integer(df$participants)
df$posters <- as.integer(df$posters)
df$orals <- as.integer(df$orals)
df$year <- as.integer(df$year)



p2012 <- subset(df, year == 2012)
pt <- paste0("AGU Conference: Number of Participants over the years.\n In 2012:", " n = ", p2012[1,"orals"], " talks, ", p2012[1,"posters"], " posters.")
pl <- ggplot(data=df, aes(x = year, y =participants, colour=conference))
pl <- pl +    scale_colour_discrete(guide = FALSE) 
pl <- pl  + geom_point( size=3)
pl <- pl  + geom_line()   # this is the base plot
pl <- pl  + geom_point(data=p2012, colour="black")   # this adds a black point
pl <- pl + geom_text(data=p2012, label=paste0("AGU 2012"), colour="black", hjust=1.1) # this adds a label for the single point

pl <- pl + ggtitle(pt)
#options(warn=w)
#qplot(year, participants , data=df, color=conference)
# with(p2.tmp, {
# qplot(participants ~ year)
# }

```
```{r echo=FALSE, fig.align='center', fig.height=3, fig.width = 5}
pl


```





--- 

## Slide 5

- Only a small subset of conference contributions, grouped by topic, are made queryable in the UI of the app.

- Data was gathered by manual screen-scraping 
- Tidying up data took much more  time than creating the shiny app.